---
title: "Case Study 1"
author: "Lani Lewis"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Here is my Zoom Presentation: https://smu-2u-com.zoom.us/rec/share/13htVJKVYg_fleqFMOzG7q1yBBF2q4hG9XXHi28xw86szepkh5uJry0yspJW6hZb.0aoM7CStP0CVqCVb
--------
This was my first group assignment which was a Budweiser EDA.



This area is used to setup R studio for this project
```{r include=FALSE}
#install.packages("caret")
#install.packages("lubridate")
#install.packages("tidyverse")


# LIBRARIES ####
library(maps) # lat and long
library(usmap)
library(caret) # confusion Matrix
library(e1071) # navebayes
library(class) # knn
#library(tm) #text mining library provides the stopwords() function
library(tidyr)
#library(plyr)
#library(jsonlite)
library(dplyr)
library(tidyverse)
library(stringr) # string manipulation
#library(mvtnorm)
library(ggplot2)
library(ggthemes) # Plot Themes
library(plotly)
```


Import in all datasets
```{r}
# IMPORT DATASETS ####
# Import Beers data set
beers <- read.csv(file = "Beers.csv",header = T)
```

```{r}
# Import Breweries data set
breweries <- read.csv("Breweries.csv",header = T)
```

```{r}
# Import Breweries data set
state_abb_list <- read.csv("State_Abbrv_List.csv",header = T)

# Change column names
colnames(state_abb_list)[1] <- c("Full_State")
colnames(state_abb_list)[3] <- c("State")
```

```{r}
# UNDERSTAND DATASETS ####
# What kind of data am I looking at
str(beers)

str(breweries)

str(state_abb_list)
```

```{r}
# REVIEW MISSING VALUES ####
# beers dataset has missing values in IBU & ABV column
view(beers)

## SFAE BEER DATASET CHECKPOINT ####
safe_beers = beers

# There is missing values in some of the rows not all. 
```


Handling Missing Data Files for Question #3
```{r}
## Look at only NA Values in beers data set ####
beers_na = beers

# create a logical vector indicating which rows have missing values
na_rows <- apply(is.na(beers_na), 1, any)

beers_na <- beers_na[na_rows, ]

str(beers_na)
```

```{r}
#### Check and make sure I only see na in ABV(3) and IBU(4) columns ####
#which(is.na(beers_na), arr.ind = TRUE)
```

```{r}
#### Count how many objects are NA in the ABV column ####
ABV_na_Index <- which(is.na(beers_na$ABV), arr.ind = TRUE)
# 62 rows

ABV_CNT <- length(ABV_na_Index)

ABV_CNT
```

```{r}
#### Count how many objects are NA in the IBU column ####
IBU_na_Index <- which(is.na(beers_na$IBU), arr.ind = TRUE)
# 1005 rows

IBU_Cnt <- length(IBU_na_Index)

IBU_Cnt
```
Investigation Code Chunk
```{r}
#### Understand which rows are missing both ABV and IBU ####
#str(which(is.na(beers_na$IBU) & is.na(beers_na$ABV), arr.ind = TRUE))

#which(is.na(beers_na$IBU) & is.na(beers_na$ABV), arr.ind = TRUE)

# 62 matches so for every ABV NA there will also be an NA IBU
# Check this theory
#beers_na[112, ] # Double Play Pilsner | 1541
#beers_na[117, ] # N Street Drive-In 50th Anniversary IPA |  1025
#beers_na[123, ] # Professor Black | 2490
#beers_na[124, ] # Little Boss | 2489
#beers_na[125, ] # Van Dayum! | 2488
```


Plot to show the missing values in correlation to the size of beer, and we see that the smaller beers have more missing values
```{r}
# MIssing Data PLot 2 ####
beers_na %>%
  ggplot() +
  geom_point(aes(y = as.factor(Style), x = as.factor(Ounces), color = as.factor(Ounces)), position = "jitter") + 
  xlab("Ounces of Beer") +
  ylab("Style of Beer") +
  labs(color = "Ounces") 
```


```{r}
# EXPORT NA FIELDS TO CSV ####
write.csv(beers_na, "Missing_Data.csv", row.names = T)
```


Merge Datasets together for Question #2
```{r}
# MERGE SAFE_BEERS AND ALL BEER DATA ####
# Rename Column Names for Consistency and Merge Process
colnames(beers)[5] <- c("Brew_ID")
colnames(beers)[1] <- c("Beer_Name")
colnames(breweries)[2] <- c("Brewery")

colnames(safe_beers)[5] <- c("Brew_ID")
colnames(safe_beers)[1] <- c("Beer_Name")


# Merge Data
#beer_ds <- merge(beers,breweries, by = "Brew_ID")

## CREATE ALL_BEER_DS ####
all_beer_ds <- merge(safe_beers,breweries, by = "Brew_ID")

# Check Data
#str(beer_ds)
str(all_beer_ds)
```


Handle Missing Values for Question #3


Replace ABV and IBU missing values with Styled median values. This way when we go to grab the median or results wont be so skewed. Especially since we have close to half of the data set missing values.
```{r}
# HANDLING MISSING VALUES WITH CUMPUTED VALUES ####
all_beer_ds$Style = as.factor(all_beer_ds$Style)

# Grab the index of each NA row by ABV and IBU ####
index_safe_beers_ABV = which(is.na(all_beer_ds$ABV))
#index_safe_beers_IBU = which(is.na(all_beer_ds$IBU))

## FInd Median based on Styel ####
style_ABV_cnt <- aggregate(ABV ~ Style, all_beer_ds, median)
style_IBU_cnt <- aggregate(IBU ~ Style, all_beer_ds, median)
```

```{r}
# Style with no IBU at all like Rauchier and replace with median first then run the loop 
missing_IBU <- setdiff(levels(all_beer_ds$Style), levels(as.factor(as.character(style_IBU_cnt$Style))))

## Median IBU for all beers ####
no_style_IBU_cnt <- median(style_IBU_cnt$IBU)


# Capture beer index for Loop
Ale_index <- which(grepl("Blonde Ale", all_beer_ds$Style))
Red_index <- which(grepl("Red Ale", all_beer_ds$Style))
APA_index <- which(grepl("American Pale Ale", all_beer_ds$Style))
Wheat_index <- which(grepl("Wheat Ale", all_beer_ds$Style))
lager_index <- which(grepl("Lager", all_beer_ds$Style))
Stout_index <- which(grepl("Stout", all_beer_ds$Style))
Mead_index <- which(grepl("Mead", all_beer_ds$Style))
Cider_index <- which(grepl("Cider", all_beer_ds$Style))
IPA_index <- which(grepl("American IPA", all_beer_ds$Style))
Double_index <- which(grepl("American Double", all_beer_ds$Style))
missing_index_IBU <- which(is.na(all_beer_ds$IBU))



# American Blonde Ale 
for (a in 1:length(Ale_index)) {
  
  all_beer_ds[Ale_index[a], "IBU"] = 15
}  


# American Amber / Red Ale
for (r in 1:length(Red_index)) {
  
  all_beer_ds[Red_index[r], "IBU"] = 17
}  


# American Pale Ale (APA)
for (p in 1:length(APA_index)) {
  
  all_beer_ds[APA_index[p], "IBU"] = 35
}  


# American Pale Wheat Ale
for (w in 1:length(Wheat_index)) {
  
  all_beer_ds[Wheat_index[w], "IBU"] = 10
}  


# lager Replacement
for (l in 1:length(lager_index)) {
  
  all_beer_ds[lager_index[l], "IBU"] = 8
}  


# Stout Replacement
for (s in 1:length(Stout_index)) {
  
  all_beer_ds[Stout_index[s], "IBU"] = 25
} 


# Mead Replacement
for (m in 1:length(Mead_index)) {
  
  all_beer_ds[Mead_index[m], "IBU"] = 0
} 


# Cider Replacement
for (c in 1:length(Cider_index)) {
  
  all_beer_ds[Cider_index[c], "IBU"] = 5
} 


# IPA Replacement
for (p in 1:length(IPA_index)) {
  
  all_beer_ds[IPA_index[p], "IBU"] = 55
} 


# Double Replacement
for (d in 1:length(Double_index)) {
  
  all_beer_ds[Double_index[d], "IBU"] = 65
} 


```

```{r}
# Catch the leftovers IBU Values
for (g in 1:length(missing_index_IBU)) {
      
  # General Replacement
  all_beer_ds[missing_index_IBU[g], "IBU"] = no_style_IBU_cnt
 
}
```

```{r}
## ABV LOOP MEDIAN ####
for (i in 1:length(index_safe_beers_ABV)) {
  all_beer_ds[index_safe_beers_ABV[i], "ABV"] = style_ABV_cnt[style_ABV_cnt$Style == all_beer_ds[index_safe_beers_ABV[i], "Style"], "ABV"]
}
```



```{r}
## Beers_ds CHECKPOINT ####
beers_ds <- all_beer_ds

str(beers_ds)
```



Compute the median alcohol content and international bitterness unit for each state. Question #4
```{r}
# DATAFRAME FOR MEDIAN AND MEAN VALUES ####
# WILL BE USED FOR PLOTTING
plot_ds <- as.data.frame(matrix(nrow = 50, ncol = 0))

plot_ds
```

```{r}
# Compute the median alcohol content and international bitterness unit for each state. ####
## IBU Median by State ####
plot_ds <- beers_ds %>% group_by(State) %>%
  summarise(median = median(IBU))

plot_ds <- arrange(plot_ds, desc(median))

colnames(plot_ds)[2] <- c("IBU_Median")

plot_ds
```

```{r}
## ABV median by State ####
ABV_Median <- beers_ds %>% group_by(State) %>%
  summarise(median = median(ABV))

ABV_Median <- arrange(ABV_Median, desc(median))

ABV_Median

# Merge data with Plot data
plot_ds <- merge(plot_ds, ABV_Median, by = "State")

colnames(plot_ds)[3] <- c("ABV_Median")

plot_ds
```


```{r}
# MERGE MEDIAN DATASET WITH STATE ABBREVIATIONS ####
plt_ds = plot_ds

# Change from factor to character
plt_ds$State <- as.character(plt_ds$State)

# Remove spaces in state field
plt_ds$State <- gsub("\\s", "", plt_ds$State)

# MERGE median values
#all_beer_ds <- merge(all_beer_ds, plt_ds, by = "State")

# Left Merge the data
plt_ds <- merge(plt_ds, state_abb_list, by = "State", all.x = T )

str(plt_ds)

```


```{r}
# MERGE AND CREATE FULL DATASET FOR CHARTS ####
## ALL BEERS DS BACKUP CHECK POINT ####
backup_ds = all_beer_ds

# RECOVER DATA
#all_beer_ds = backup_ds 

# Remove spaces in state field
#all_beer_ds$State <- gsub("\\s", "", all_beer_ds$State)

#all_beer_ds <- merge(all_beer_ds, state_abb_list, by = "State", all.x = T )

#str(beer_ds)
str(all_beer_ds)
str(backup_ds)
```



Compute the Number of Breweries by State Question #1
```{r}
# How many breweries are present in each state? | State_cnt ####
state_cnt <- all_beer_ds %>% count(State, sort = T)

colnames(state_cnt)[2] <- "Number_of_States"

state_cnt <- arrange(state_cnt, desc(Number_of_States))

# Remove spaces in state field
state_cnt$State <- gsub("\\s", "", state_cnt$State)

# Left Merge 
state_cnt = merge(state_cnt, state_abb_list, by = "State", all.y = T)


str(state_cnt)
```


```{r}
# PLOT1 Facet Wrap| How many breweries are present in each state? ####
state_cnt %>% arrange(state_cnt, desc(Number_of_States)) %>%
  ggplot(aes(x = Number_of_States, fill = Full_State)) +
  geom_histogram() +
  xlab("Number of Breweries by States") +
  ylab("") +
  facet_wrap(Number_of_States ~ Full_State) +
  scale_y_discrete() +
  theme(legend.position = "none")
  
```

```{r}
# PLOT2 Bar Chart| How many breweries are present in each state? ####
Top_Ten <- state_cnt[order(-state_cnt$Number_of_States), ]
  
head(Top_Ten, 10) %>% 
  ggplot(aes(reorder(Number_of_States, x = Full_State), x = Number_of_States, fill = Full_State, color = "red")) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Number_of_States), size = 4, hjust = 1, color = "black") + 
  xlab("Number of Breweries") +
  ylab("State") +
  theme(legend.position = "none") +
  scale_fill_brewer(palette = "YlOrRd")




# ?color
```


Median Charts for Question #4

Bar Charts to compare the Median value of ABV vs IBU by state
```{r}
Top_5_IBU <- plt_ds[order(-plt_ds$IBU_Median), ]
Top_5_ABV <- plt_ds[order(-plt_ds$ABV_Median), ]
  

# IBU CHART
p3 <- head(Top_5_IBU, 5) %>% 
  ggplot(aes(reorder(IBU_Median, x = Full_State), x = IBU_Median, fill = IBU_Median)) +
  geom_bar(stat = "identity") +
  xlab("IBU Median") +
  ylab("State")  +
  scale_fill_gradient(low = "black", high = "red") 
  


# ABV CHART
p4 <- head(Top_5_ABV, 5) %>% 
  ggplot(aes(reorder(ABV_Median, x = Full_State), x = ABV_Median, fill = ABV_Median)) +
  geom_bar(stat = "identity") +
  xlab("ABV Median") +
  ylab("State") +
  scale_fill_gradient(low = "black", high = "red")




cowplot::plot_grid(p3, p4, ncol = 2)
```




Bar Charts to compare the TOP 5 State and their Median value of ABV vs IBU
```{r}
# IBU CHART
p1 <- plt_ds %>% ggplot(aes(reorder(IBU_Median, x = State), x = IBU_Median, fill = IBU_Median)) +
  geom_bar(stat = "identity") +
  xlab("IBU Median") +
  ylab("State")  +
  scale_fill_gradient(low = "black", high = "red") +
  theme(axis.text.y = element_blank()) 
  


# ABV CHART
p2 <- plt_ds %>% ggplot(aes(reorder(ABV_Median, x = State), x = ABV_Median, fill = ABV_Median)) +
  geom_bar(stat = "identity") +
  xlab("ABV Median") +
  ylab("State") +
  scale_fill_gradient(low = "black", high = "red")  +
  theme(axis.text.y = element_blank())




cowplot::plot_grid(p1, p2, ncol = 2)
```


Breakdown of of Median values by State
```{r}
# PLOT IBU Median by States ####
plt_ds %>% arrange(IBU_Median, desc(Full_State)) %>%
  ggplot(aes(x = IBU_Median, fill = Full_State)) +
  geom_histogram() +
  xlab("IBU Median by States") +
  ylab("") +
  facet_wrap(IBU_Median ~ Full_State) +
  scale_y_discrete() +
  theme(legend.position = "none")

```



```{r}
# FIX ROUNDING ####
plt_ds$ABV_Median <- round(plt_ds$ABV_Median, digits = 3)

# PLOT ABV Median by States ####
plt_ds %>% arrange(ABV_Median, desc(Full_State)) %>%
  ggplot(aes(x = ABV_Median, fill = Full_State)) +
  geom_histogram() +
  xlab("ABV Median by States") +
  ylab("") +
  facet_wrap(ABV_Median ~ Full_State) +
  scale_y_discrete() +
  theme(legend.position = "none")

```
Review the Summary Data of the data sets for Question #6
```{r}
# Reviewing all beers ds | The entire data set
summary(all_beer_ds)

```

```{r}
# Reviewing plt ds | Median Values of the data set
summary(plt_ds)

```



Which state has the maximum alcoholic (ABV) beer? Question #5
```{r}
# Max ABV State
Max_ABV <- head(Top_5_ABV, 1)
Max_ABV
```


```{r}
# MAX IBU State
Max_IBU <- head(Top_5_IBU, 1)
Max_IBU
```


```{r}
median_map <- plt_ds

colnames(median_map)[1] <- c("state")
colnames(Max_ABV)[1] <- c("state")
colnames(Max_IBU)[1] <- c("state")



plot_usmap(data = median_map, values = "ABV_Median") +
  labs(title = "Median Heatmap by State") +
  scale_fill_gradient(low = "dark grey", high = "red") 

```
  
```{r}
# Kentucky Map ABV MAX STATE
plot_usmap(data = median_map, values = "ABV_Median", include = c("KY"),  labels = TRUE, label_color = "white", show.legend = FALSE) +
  labs(title = paste(Max_ABV$Full_State, " Max ABV ", Max_ABV$ABV_Median)) +
  scale_fill_gradient(low = "dark grey", high = "red")
```

```{r}
plot_usmap(data = median_map, values = "IBU_Median") +
  labs(title = "Median Heatmap by State") +
  scale_fill_gradient(low = "dark grey", high = "red") 
```

```{r}
# Kentucky Map ABV MAX STATE
plot_usmap(data = median_map, values = "ABV_Median", include = c("NJ"),  labels = TRUE, label_color = "white", show.legend = FALSE) +
  labs(title = paste(Max_IBU$Full_State, " Max IBU ", Max_IBU$IBU_Median)) +
  scale_fill_gradient(low = "dark grey", high = "red")
```

Question #7: Is there an apparent relationship between the bitterness of the beer and its alcoholic content? 
```{r}
# MEDIAN PLOT | relationship between IBU and ABV ####
plt_ds %>% ggplot() +
  geom_point(aes(x = ABV_Median , y = IBU_Median, color = ABV_Median, alpha = IBU_Median)) +
  geom_smooth(aes(x = ABV_Median , y = IBU_Median), color = "red") +
  scale_color_gradient2(low = "black", high = "red") +
  theme(legend.position = "none")
  
```

```{r}
# All Data PLOT | relationship between IBU and ABV ####
all_beer_ds %>% ggplot() +
  geom_point(aes(x = ABV, y = IBU, color = ABV), position = "jitter") +
  geom_smooth(aes(x = ABV, y = IBU, color = IBU), color = "red") +
  scale_color_gradient2(low = "black", high = "red")  +
  theme(legend.position = "none")
  
```

What is the probability of picking a beer like Budweiser lager in this list of craft beers.
```{r}
# KNN Classification | FOR PRESENTATION ####
train = all_beer_ds

test = data.frame(ABV = .05, IBU = .12) 

#all_beer_ds[, 5:6]
class <- knn(train[, 4:5], test, train$Style, k = 5, prob = T)

class

#table(class, test)
```


Question #8 - difference with respect to IBU and ABV between IPAs (India Pale Ales) and other types of Ale (any beer with “Ale” in its name other than IPA). 
```{r}
# Create KNN Train and Test Datasets ####

train2 = train[grepl("Ale|IPA", train$Style), ] # Everything Ale and IPA

train2$Classification = ifelse(grepl("\\bAle\\b", train2$Style), "Ale", "IPA" )

# ------------------------ #

train3 = train2[!grepl("\\bIPA\\b", train2$Style), ] # ONLY ALE 

test2 = train2[grepl("\\bIPA\\b", train2$Style), ] # ONLY IPA

#str(train2)

#str(train3)

#str(test2)
```



What is the probability that we will find similar style beers within the entire dataset based off the ABV and IBU
```{r}
class4 <- knn.cv(train2[, c(4:5)], train2$Classification, k = 5, prob = T)

# table(class5, train2$Style)

confusionMatrix(table(class4, train2$Classification))
```





